%%!TEX root = ./UserManual.tex
\chapter{Concepts and Algorithms}
\label{chap:code}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% GenGeoPop
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{GenGeoPop}
\label{section:gengeopop}

\subsection{Background}
\label{subsection:background}

To explain the algorithms used for generating the geography of the countries and their respective population, we have to introduce some background concepts:

\begin{description}
    \item[ContactPool]:
        A pool of persons that potentially come contacts with each other that potentially lead to disease transmission.
        We distinguish different types of ContactPools associated with the workplace, school, community and households.
        Note that there is a difference between the size of a school (for example 500 students) and the size of the ContactPools inside the school (20 students taking their classes together all the time).
    \item[ContactCenter]:
        A group of one or more ContactPools of the same type.
        This allows for a school or College to contain multiple classes (ContactCenter).
        A ContactCenter may contain a single ContactPool, for example for a Household.
    \item[K-12 student]: 
    		Persons from 3 until 18 years of age that are required (at least in Belgium) to attend school. 
    		Students that skip or repeat years are not accounted for.
    \item[College student]:
        Persons older than 18 and younger than 26 years of age that attend an institution of higher education. 
        For simplicity we group all forms of higher eduction into the same type of ContactCenter, a College. 
        A fraction of college students will attend a college ``close to home'' and the others will attend a 
        college ``far from home''. Most higher educations don't last 8 years, but this way we compensate 
        for changes in the field of study, doctoral studies, advanced masters and repeating a failed year of study.
    \item[Employable]:
        We consider people of ages 18 to 65 to be potentially employable. A fraction of people between 18 and 26 
        will attend a college, and the complementary fraction will be employable.
    \item[Active population]:
        The fraction of the employable population that is actually working. A fraction of these workers will 
        work ``close to home'' and the complementary fraction will commute to a workplace ``far from home''.
    \item[Household profile]:
        The composition of households in terms of the number of members and their age is an important 
        factor in the simulation. In this case the profile is not defined by the age of its members or fractions, 
        but through a set of reference households. This set contains a sample of households which is 
        representative of the whole population in their composition.
    \item[GeoGrid locations]:
        Our data only allows for a limited geographical resolution. We have the longitude and latitude of cities and municipalities (a distinction we will not make), which we shall use to create GeoGrid locations for the domicile of the households. All households in the same location are mapped to the coordinates of the location's center. These locations with corresponding coordinates will form a grid that covers the simulation area.
\end{description}

\subsection{Generating the geography}
\label{subsection:gengeo}
We start by generating the geographical component, a GeoGrid.
It contains locations with an id, name, province, coordinates and a reference population count.

The latter requires some comment. When we build a synthetic population, we start by generating households. That household is the assigned a location by drawing from a discrete distribution of locations with weights proportional to the  relative population count of the location. As a consequence, each location will have a population count that differs stochastically from the reference count, but will be close to it. When many synthetic populations are generated, the average of a location 's population count will tend to its reference count.

Locations contain multiple ContactCenters, like Schools and Households, which in turn contain ContactPools.
This structure is internally generated by several ``Generators" and will afterwards be used by ``Populators" to fill the ContactPools. The different types of ContactCenters are created by a different partial generator for each type and added separately to the GeoGrid. We construct the following types of ContactCenters:

\begin{description}
    \item[Households]:
        The number of households is determined by the average size of a household in the reference 
        profile and the total population count.
        The generated households are then assigned to a location by a draw from a discrete distribution of 
        locations with weights proportional to the  relative population count of the location.
    \item[K-12 Schools]:
        Schools a reference count of 500 students, with a refernce count of 20 students per class,
         corresponding to 25 ContactPools per school. 
         The total number of schools in the region is determined by the population count, the fraction 
         of people in the K-12 school age bracket and of course the reference school size.
        The algorithm for assignment of schools to locations is similar to that of households.
    \item[Colleges]:
        Colleges have a reference count of 3000 students with 150 students per ContactPool.
        Colleges are exclusively assigned to the 10 locations with highest reference population count (cities).
        For these 10 locations we use a discrete distribution with weights proportional to the population count 
        of the city relative to the total population of the 10 cities.
    \item[Workplaces]:
        The algorithm for assignment of workplaces to locations is analogous to that of households, 
        but here we factor in commuting information to determine the actual number of workers at a location.
        That is, the reference count of active persons (i.e. persons assigned to a workplace) is 
        the number of active persons living at a location (i.e. their household is located there)
        plus the active persons commuting out of that location minus the number of active persons
        commuting into that location. 
        The reference count for persons at a workplace is 20.
    \item[Communities]:
        We create both primary and secondary communities. Each has a  reference count of 2000 persons. 
        Communities consist of persons from all ages. 
        The assignment to locations is again similar to that of households.
\end{description}

\subsection{Generating the population}
\label{subsection:genpop}
After creating the structures that will allow people to come in contact with each other, we have create the population itself and determine the different ContactPools they will be in. The persons are created based on the Household profiles in the HouseholdPopulator. Similar to the approach the ContactCenters, we have a partial populators that 
each populates ia type of ContactPool:

\begin{description}
    \item[Households]:
        To create the actual persons, we randomly draw a household from the list of reference households 
        and use that as a template for the number of household members and their ages.
        We simply do this for each household in the GeoGrid, since we already determined 
        the locations and number of households while generating the geography.
    \item[K-12 Schools]:
        We start by listing all schools within a 10km radius of the household location. 
        If this list is empty, we double the search radius until it's no longer empty.
        We then randomly select a ContactPool from those in the schools of the list,
       even if this ContactPool now has more students than average.
    \item[Colleges]:
        Students who study ``close to home" are assigned to a college with an algorithm
        similar to the assignment to K-12 schools.
        Students that study ``far from home" we first determine the list of locations 
        where people from this locations commute to. We randomly select one of these locations 
        using a discrete distribution based on the relative commuting information.
        We then randomly choose a ContactPool at a college in this location and assign it to 
        the commuting student.
    \item[Workplaces]:
        We first decide whether the person is active (correct age, fraction of the age bracket that is a student,
        fraction of the age bracket that is active).
        We assign a workplace to an active person that works ``close to home" in an 
        analogous way as the assignment of K-12 schools to students.
        For the commuting workers we use an algorithm analogous to that of the commuting college students.
    \item[Communities]:
        The communities we can choose from at a location is determined in an analogous 
        way to the K-12 schools.
        For primary communities we randomly select, for each person in a household, a ContactPool 
        from the list of Communities within a 10km radius. If the list empty, \ldots see the K-12 algoritm.
        In secondary communities however, we assign complete households to the 
        ContactPools instead of each person in the household separately.
\end{description}
